---
title: "RAAC in English schools - 2023-12-06"
format: 
  html:
    toc: true
author:
  - name: Peter EJ Kemp
    email: peter.kemp@kcl.ac.uk
    affiliations:
    - name: King's College London
number-sections: true
editor: source
tbl-cap-location: top
---

```{r}
#| echo: false
#| warning: false

#setup_TRACER_vars()
#?desc


# TODO: find out why 139218 149972 aren't in the schools df, update schools to tell us the type of school, e.g. primary, secondary, 16 plus etc

#| warning: false
#| echo: false
node <- Sys.info()

if(node["nodename"] == "KCL8KWNXY2"){
  # basefolder <- r"(C:\Users\w1926273\OneDrive - King's College London\Code\SCARI\)"
  basefolder <- r"(C:/Users/w1926273/Google Drive/Kings/Code/PISR/)"
}else if(node["nodename"] == "PETES_PC"){
  basefolder <- r"(C:/Users/Peter/Google Drive/Kings/Code/PISR/)"
  old_proj_folder <- r"(C:/Users/Peter/Google Drive/Kings/Code/OpenAnalysis/)"
}else if(node["nodename"] == "KCL3Y7NKR3"){ # new laptop
  basefolder <- r"(G:\My Drive\Kings\Code\PISR\)"
  old_proj_folder <- r"(C:/Users/Peter/Google Drive/Kings/Code/OpenAnalysis/)"
}

library(tidyverse)
library(readODS)
library(gt)
library(glue)

data_folder <- r"[C:\Users\Peter\Google Drive\Kings\Code\PISR\Data\]"
# data_folder <- r"[D:\Data\EdR\]"

source(paste0(basefolder, "_reports/TRACER/TRACER-setup.R"))
#source(paste0(codefolder, 'OpenAnalysis_load.R'))

library(janitor)

# load the dataset (with crazy skip)
# raac_raw <- read_ods(glue(basefolder, "_reports/RAAC/RAAC.ods"), skip=21) %>% 
raac_raw <- read_excel(glue(basefolder, "_reports/RAAC/RAAC.xlsx"), 
                       skip=23, 
                       sheet="List of settings") %>% 
  clean_names() %>%
  filter(`setting_name` != "Not applicable") %>%
  mutate(urn = as.numeric(urn)) %>%
  dplyr::rename("setting_mitigation" = "setting_mitigation_27_11_23")

raac <- left_join(raac_raw, 
          schools %>% 
            group_by(URN) %>% 
            filter(year == max(year)),
          by=c("urn"="URN"))

raac <- raac %>% 
  mutate(lbl = paste("<p>",as.character(`setting_name`), "</p><p>", `setting_phase`, "</p>", `setting_mitigation`, "</p>"),
         details = paste("<p><b>",`setting_name`, "</b></p>",
                              " <a href='https://www.get-information-schools.service.gov.uk/Establishments/Establishment/Details/",
                             urn,"'>",urn,
                         "</a></br> <a href='https://", `setting_website`, "'>website</a>",
                         "<table>",
                         "<tr><td>phase:</td><td>", `setting_phase`, "</td></tr>",
                         "<tr><td>size:</td><td>", stu_total, "</td></tr>",
                         "<tr><td>mitigation:</td><td>", `setting_mitigation`, "</td></tr>",
                         "<tr><td>LA:</td><td>", LA, "</td></tr>",
                         "<tr><td>Region:</td><td>", Region, "</td></tr>",
                         "</table>")) %>%
  mutate(clr = case_when(
    `setting_phase` == "Primary" ~ "#97EFE9",
    `setting_phase` == "Secondary" ~ "#4A001F",
    `setting_phase` == "16 plus" ~ "Red",
    `setting_phase` == "Not applicable" ~ "#000000",
    `setting_phase` == "All-through" ~ "#A7C4C2" 
  ))
                         

                         # "<div><span style='width: 50%;'>phase:</span>", `Setting Phase`, "</div>",
                         # "<div><span style='width: 50%;'>size:</span>", stu_total, "</div>",
                         # "<div><span style='width: 50%;'>mitigation:</span>", `Setting mitigation`, "</div>",
                         # "<div><span style='width: 50%;'>la:</span>", LA, "</div>"))

### leaflet
library(leaflet)
library(sf)
library(scales)
library(htmltools)

# need to fix Lat Lon for schools


#get northing easting
edubase <- read_csv("D:\\Data\\EdR\\Edubase\\edubasealldata.csv")
raac <- left_join(raac %>% select(-Easting, -Northing), 
                  edubase %>% 
            select(URN, Easting, Northing) %>% 
            filter(URN %in% raac$urn) %>% distinct(),
          by=c("urn"="URN")) %>% distinct() 
  

lat_lon <- raac %>%
  st_as_sf(coords = c("Easting", "Northing"), crs = 27700) %>%
  st_transform(4326) %>%
  st_coordinates() %>%
  as_tibble()


raac$lat <- lat_lon$Y
raac$lon <- lat_lon$X

# set colour scale
pal <- colorNumeric(c("black", "green"), 
                    domain=c(min(raac$stu_total, na.rm = TRUE):max(raac$stu_total, na.rm = TRUE)))

map_raac_provision <- leaflet(data=raac) %>%
  # addProviderTiles("Stamen.Toner") %>%
  addTiles() %>%
  addCircleMarkers(~lon, ~lat, 
                   fillColor = ~clr, # ~pal(stu_total), 
                   #fillColor = "black", 
                   color = "black", 
                   weight =1, opacity = 1, fillOpacity = 1,
                   popup = ~details, 
                   radius = ~rescale(stu_total, c(1,10)) + 3.5,
                   label = ~map(lbl, htmltools::HTML))
  # addLegend("bottomright", 
  #           pal = pal, 
  #           values = ~Geography,
  #           title = "% taking Geography",
  #           opacity = 1)

```

# Introduction

A very quick lunchtime report on school providers with RAAC and their mitigation plans. Data from [here](https://www.gov.uk/government/publications/reinforced-autoclaved-aerated-concrete-raac-management-information) and up to date as of 2023-12-06, using the DfE dataset that states: "List of education settings with confirmed RAAC as of **27 November 2023**". Data likely to change, mistakes highly likely mine, please drop me an email if you have any suggestions / questions.

-   summary tables are for state schools and state school students
-   Total student figures come from the 2021/22 DfE dataset
-   it's possible schools using new URNs as of 2022/23 or 2023/24 will not be included in this list (e.g. academy converters)
-   The return rate for each LA is still unclear. Low occurrence of RAAC might be due to high return rate. This needs clearing up.
-   The DfE have changed the format of the data file again, from ods to xlsx, with two sheets rather than one. Some errors might appear through that.

# Map of providers with RAAC

-   Size of dots: number of students in provider
-   Colours
    -   Man City Blue: Primary;
    -   Wine: Secondary;
    -   Grey: All through;
    -   Man United Red: 16 plus;
    -   Black: Not applicable;

```{r}
#| warning: false
#| echo: false
map_raac_provision
```

```{r}
#| echo: false
#| warning: false
raac_all <- schools %>% 
            group_by(URN) %>% 
            filter(year == 2023,
                   schType %in% c("Comprehensive", "Special", "Grammar")) %>%
            mutate(raac = URN %in% raac$urn)

# find missing URNs
# raac$URN
# final_URN <- raac_all %>% filter(raac) %>% pull(URN)
# setdiff(raac$URN, final_URN)
# 
# # Missing URNS
# # because they weren't entering GCSEs in 2022? Or new schools
# https://get-information-schools.service.gov.uk/Establishments/Establishment/Details/139218
# https://get-information-schools.service.gov.uk/Establishments/Establishment/Details/149972


overall_stu <- raac_all %>% 
  group_by(raac) %>% 
  dplyr::summarise(students = sum(stu_total, na.rm=TRUE)) %>%
  pivot_wider(names_from = raac, values_from = students) %>%
  dplyr::rename(student_no = `FALSE`,
         student_yes = `TRUE`) %>%
  mutate(`student_%` = 1*(student_yes / (student_no + student_yes)))

overall_prov <- raac_all %>% 
  group_by(raac) %>% 
  dplyr::summarise(providers = n()) %>%
  pivot_wider(names_from = raac, values_from = providers) %>%
  dplyr::rename(provider_no = `FALSE`,
         provider_yes = `TRUE`) %>%
  mutate(`provider_%` = 1*(provider_yes / (provider_no + provider_yes)))

overall_all <- cbind(overall_stu, overall_prov)

# LA

la_stu <- raac_all %>% 
  group_by(LA, raac) %>% 
  dplyr::summarise(students = sum(stu_total, na.rm=TRUE)) %>%
  pivot_wider(names_from = raac, values_from = students) %>%
  dplyr::rename(student_no = `FALSE`,
         student_yes = `TRUE`) %>%
  mutate(`student_%` = 1*(student_yes / (student_no + student_yes)))

la_prov <- raac_all %>% 
  group_by(LA, raac) %>% 
  dplyr::summarise(providers = n()) %>%
  pivot_wider(names_from = raac, values_from = providers) %>%
  dplyr::rename(provider_no = `FALSE`,
         provider_yes = `TRUE`) %>%
  mutate(`provider_%` = 1*(provider_yes / (provider_no + provider_yes)))

la_all <- left_join(la_stu, la_prov)

# Missing schools
sch_missing <- setdiff(raac$urn, raac_all$URN)
sch_missing <- raac %>% 
  filter(urn %in% sch_missing) %>% 
  select(urn, setting_name, local_authority, type_of_setting) %>% 
  mutate(urn = glue('<p><a href = \"https://www.get-information-schools.service.gov.uk/Establishments/Establishment/Details/{urn}\">{urn}</a>'),
         urn = map(urn, gt::html)) %>% gt()

# number of children in each la in raac schools
# percentage of children in each La in raac schools
# number of providers in each la with raac
# percentage of providers in each la with raac

##### REGION

region_stu <- raac_all %>% 
  group_by(Region, raac) %>% 
  dplyr::summarise(students = sum(stu_total, na.rm=TRUE)) %>%
  pivot_wider(names_from = raac, values_from = students) %>%
  dplyr::rename(student_no = `FALSE`,
         student_yes = `TRUE`) %>%
  mutate(`student_%` = 1*(student_yes / (student_no + student_yes)))

region_prov <- raac_all %>% 
  group_by(Region, raac) %>% 
  dplyr::summarise(providers = n()) %>%
  pivot_wider(names_from = raac, values_from = providers) %>%
  dplyr::rename(provider_no = `FALSE`,
         provider_yes = `TRUE`) %>%
  mutate(`provider_%` = 1*(provider_yes / (provider_no + provider_yes)))
  
region_all <- left_join(region_stu, region_prov)

```

# Current distribution of providers and students

```{r overall}
#| warning: false
#| echo: false
overall_all %>% ungroup() %>% gt() %>% 
  sub_missing(missing_text = 0) %>%
  fmt_percent(columns = c("student_%", "provider_%"),
              decimals = 2) %>%
  cols_label(
    student_no = "Total students",
    student_yes = "Students in RAAC",
    `student_%`	= "% Students in RAAC",
    provider_no = "Total providers",
    provider_yes = "Providers with RAAC",
    `provider_%` = "% Providers with RAAC")
  
```

Some schools with RAAC are missing from this dataset:

```{r overall_by_school_type}
#| warning: false
#| echo: false
sch_missing


# get info on total % of primary and secondary schools closed

```

# Current distribution of providers by Region

```{r region}
#| warning: false
#| echo: false
region_all %>% ungroup() %>% arrange(desc(`student_%`)) %>%
  gt() %>% 
  sub_missing(missing_text = 0) %>%
  fmt_percent(columns = c("student_%", "provider_%"),
              decimals = 2) %>%
  data_color(
    columns = c("student_%", "provider_%"),
    method="numeric",
    palette = c("red", "green")) %>%
    tab_spanner_delim(
    delim = "_"
  )
```

# Current distribution of providers by Local Authority

```{r LA}
#| warning: false
#| echo: false
la_all %>% ungroup() %>% arrange(desc(`student_%`)) %>%
  gt() %>% 
  sub_missing(missing_text = 0) %>%
  fmt_percent(columns = c("student_%", "provider_%"),
              decimals = 2) %>%
  data_color(
    columns = c("student_%", "provider_%"),
    method="numeric",
    palette = c("red", "green")) %>%
    tab_spanner_delim(
    delim = "_"
  ) #%>%
  #tab_options(ihtml.active = TRUE,
  #            ihtml.use_filters = TRUE)
  
  
```
